---
title: "eXtreme Gradient Boosting for Alzheimer's Disease CSF Biomarker Discovery"
author: "Justin Taylor"
date: "May 10, 2016"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: "simplex"
    highlight: "textmate"
    
---

Load the `AlzheimerDisease` data set from the AppliedPredictiveModeling package.  

```{r Dependencies, include = FALSE}
library(AppliedPredictiveModeling)
library(xgboost)
library(DT)
library(magrittr)
library(dplyr)
library(caret)
```

# Data {.tabset}  

```{r}
data(AlzheimerDisease)
```  

## Inspection  

Dimensions: `r dim(predictors)`  

Number of rows = Number of labels: `r nrow(predictors) == length(diagnosis)`   

Variable names:  
```{r}
colnames(predictors)
```

Variable types:  

```{r}
sapply(predictors, class) %>% table
```

Head of selected variables:  

```{r}
head(predictors[,1:5])
```

```{r cleanData}
ad_matrix <- as.matrix(sapply(predictors, as.numeric))
```

The `label` parameter for `xgboost` should be $0$ or $1$ for binary classification.  

```{r fixBinaryLabels}
diagnosis <- parallel::mclapply(diagnosis, function(x) (
  if (x == "Control") (0) else 1
)) %>% unlist; head(diagnosis)
```

## Preview  

```{r previewData}
DT::datatable(apply(ad_matrix, 2, round, 3), extensions = c('Buttons', 'Responsive'),
              options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))
```

## Splitting
Simple splitting with 60/40 train/test spread.  

```{r}
set.seed(1983)
train_index <- createDataPartition(diagnosis, p = .6, list = FALSE, times = 1)
```

----------------------------------------------------------------------------  


# Train Model  

```{r}
model_params <- list(objective = "binary:logistic", eta = 0.01, max.depth = 3, eval_metric = "auc")
```

```{r trainModel}
model <- xgboost::xgboost(data = ad_matrix[train_index,], label = diagnosis[train_index], 
                          max.depth = 6, eta = 1, nthread = 2, nround = 8, 
                          objective = "binary:logistic")

#model_cv <- xgb.cv(params = xgb_params, data = ad_matrix, nrounds = 100, nfold = 5, label = diagnosis)
# plot(model_cv$train.auc.mean)

```

----------------------------------------------------------------------------  


# Feature Importance {.tabset}

## Table  

```{r getFeatureImportance}
importance <- xgb.importance(colnames(predictors), model = model); 
DT::datatable(importance)
```

## Plot  

```{r plotFeatureImportance, width = 10, height = 8}
#install.packages('Ckmeans.1d.dp')
library(Ckmeans.1d.dp)
xgb.plot.importance(importance)
```

----------------------------------------------------------------------------  


# Testing Model {.tabset}

## Predictions

```{r}
predictions <- predict(model, ad_matrix[-train_index,])
predictions <- as.numeric(predictions > 0.5)
table(predictions)
table(diagnosis[-train_index])
```

## Confusion Matrix  

```{r}
confusionMatrix(predictions, diagnosis[-train_index])
```

```{r}
error <- mean(predictions != diagnosis[-train_index])
```

Test Error: `r error`

---------------------------------------------------------------------------- 

# References

Craig-Schapiro, R., Kuhn, M., Xiong, C., Pickering, E. H., Liu, J., Misko, T. P., … Holtzman, D. M. (2011). Multiplexed Immunoassay Panel Identifies Novel CSF Biomarkers for Alzheimer’s Disease Diagnosis and Prognosis. PLoS ONE, 6(4), e18850. http://doi.org/10.1371/journal.pone.0018850  

Tianqi Chen, Tong He and Michael Benesty (2016). xgboost: Extreme Gradient Boosting. R
  package version 0.4-3. https://CRAN.R-project.org/package=xgboost
  
Max Kuhn. Contributions from Jed Wing, Steve Weston, Andre Williams, Chris Keefer, Allan
  Engelhardt, Tony Cooper, Zachary Mayer, Brenton Kenkel, the R Core Team, Michael Benesty,
  Reynald Lescarbeau, Andrew Ziem, Luca Scrucca, Yuan Tang and Can Candan. (2016). caret:
  Classification and Regression Training. R package version 6.0-68.
  https://CRAN.R-project.org/package=caret
  
Max Kuhn and Kjell Johnson (2014). AppliedPredictiveModeling: Functions and Data Sets for
  'Applied Predictive Modeling'. R package version 1.1-6.
  https://CRAN.R-project.org/package=AppliedPredictiveModeling
  

---------------------------------------------------------------------------- 



# Session Info

```{r}
sessionInfo()
```


